---
title: "Assignment 7"
author: "Adriana LH"
date: "10/31/2020"
output: 
 html_document:
    theme: cosmo
    highlight: zenburn
---
```{r load libraries, include=FALSE}
library(tidycensus)
library(tidyverse)
library(sf)
library(leaflet)
library(htmltools)
library(htmlwidgets)
library(raster)
library(gstat)
library(spatial)
```

## Attaching my csv files

The dataset I'm looking at tabulates all the evictions happening in each Boston census tract over the past 43 weeks. I realize that with the covid-induced eviction moratorium that has taken place over most of this year, this data is rather atypical (at least in 2020). However, I still thought it was interesting to investigate. 

```{r}
evictions_boston <- read_csv("boston_20201024.csv") %>%
  mutate(GEOID10 = as.numeric (str_sub(GEOID,-11,-1))) 

boston_tracts <- st_read("https://bostonopendata-boston.opendata.arcgis.com/datasets/4a8eb4fb3be44ed5a1eec28551b9f3b2_0.geojson?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D") %>%
   mutate (GEOID10 = as.numeric(GEOID10))
```
##Cleaning my data

```{r}
test <- evictions_boston %>%
  group_by(GEOID10, racial_majority) %>%
summarise(total = sum(filings_2020))
```
## Choropleth Map

```{r}
evictions <- left_join(boston_tracts, test, by = "GEOID10") %>%

na.omit(evictions)

evictions$label <- 
  paste(evictions$racial_majority, "racial majority <br>", 
        evictions$total, "total evictions in each tract") %>% 
  lapply(htmltools::HTML)

bins <- seq(min(evictions$total),
            max(evictions$total), by = 5)
pal <- colorNumeric("viridis", 
                    domain = evictions$total,
                    na.color = "#00000000")

leaflet(evictions) %>%
  addProviderTiles(providers$Stamen.Toner) %>%
  addPolygons(highlightOptions = highlightOptions(fillOpacity = 1),
              label = ~label,
              fillColor = ~pal(total),
              weight = 1, color = "black") %>% 
   addControl('<a href="https://evictionlab.org/eviction-tracking/get-the-data/">Data source</a>',
             position = "topleft") %>%
  addControl("Evictions in Boston Through the first 43 Weeks of 2020", position = "bottomleft") %>%
  addLegend(pal = pal, 
            values = ~total,
            bins = 10,
            opacity = 0.75, title = "Total Evictions per Tract",
            position = "topright")

```

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs "

WGS84 <- "+proj=longlat +ellps=WGS84 +datum=WGS84 +no_defs"

tract_points <- st_centroid(
  st_transform(evictions, crs = MA_state_plane)) %>%
  st_transform(WGS84)

leaflet(tract_points) %>%
  addProviderTiles(providers$Stamen.Toner) %>%
  addCircles(label = ~label,
             fillColor = ~pal(total),
             stroke = FALSE, 
             radius = 200, 
             fillOpacity = 1) %>% 
   addControl('<a href="https://evictionlab.org/eviction-tracking/get-the-data/">Data source</a>',
             position = "topleft") %>%
  addControl("Evictions in Boston Through the first 43 Weeks of 2020", position = "bottomleft") %>%
  addLegend(pal = pal, 
            values = ~total,
            bins = 10,
            opacity = 0.7, title = "Total Evictions in Each Tract",
            position = "topright")
```

```{r}
tract_points_sp <- tract_points %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

tract_poly_sp <- evictions %>%
  st_transform(MA_state_plane) %>%
  as_Spatial()

eviction_raster <- raster(tract_poly_sp, res=10)

gs <- gstat(formula=total~1, locations=tract_points_sp)
idw_interp <- interpolate(eviction_raster, gs)

idw_interp_clip <- mask(idw_interp, tract_poly_sp)

leaflet(tract_points) %>%
  addProviderTiles(providers$Stamen.TonerLite) %>%
  addRasterImage(idw_interp_clip, colors = pal, opacity = 0.8) %>% 
   addControl('<a href="https://evictionlab.org/eviction-tracking/get-the-data/">Data source</a>',
             position = "topleft") %>%
  addControl("Evictions in Boston Through the First 43 Weeks of 2020",
             position = "bottomleft")%>%
  addLegend(pal = pal, 
            values = ~total,
            bins = 10,
            opacity = 0.7, title = "Total Evictions in Each Tract",
            position = "topright")
```
## Discussion of Map Fit/Appropriateness

This data set was unique in the sense that it had one enormous outlier (50), with the majority of the eviction numbers being between 0 and 10. This meant that a map that allows the viewer to investigate the data and understand the granularity of it themselves is going to fit the data best. For this reason, the choropleth map (map #1) displaysbest map fit. With that map, the viewer is able to more clearly see the range of evictions between census tracts.

Map #2 does also show that data spread accurately, but of course, with dozens of circles clustered on the map, it's just harder to comprehend. 

Map #3, the raster interpolation map, shows this eviction data least faithfully. Due to its nature of smoothing data subtleties, t may give the viewer the (incorrect) sense that evictions are only happening in certain hotspot locations of the city. 
