---
title: "Vis Assignment 4"
author: "Adriana LH"
date: "10/2/2020"
output:
   html_document:
    theme: cosmo
    highlight: zenburn
---

```{r load libaries, include=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(ggplot2)
library(png)
library(ggrepel)
library(ggimage)
```

## Loading Data
I am plotting the non-public schools in Boston.

```{r}
non_public <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/0046426a3e4340a6b025ad52b41be70a_1.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
```

```{r}
  opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/boston_streets.osm')
```

```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

boston_street_features <- opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

boston_streets <- boston_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)

```

## Plotting street data

```{r}
ggplot(boston_streets) +
  geom_sf() +
  theme_map()
```

## Downloading OTP, Java, etc.
```{r, message = FALSE, warning = FALSE}
path_otp <- otp_dl_jar("OTP")
```

```{r, include= FALSE}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024) 
```

```{r, message= FALSE, warning = FALSE}
otp_setup(otp = path_otp, dir = path_data, memory = 1024)
```

```{r}
otpcon <- otp_connect()
```

## Creating Isochrones
I am plotting the isochrones for a 5 min walk and 5 min bike ride.

```{r, message = FALSE, warning = FALSE}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = non_public, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")
```

```{r, message = FALSE, warning = FALSE}
iso_5min_bike <- 
  otp_isochrone(otpcon = otpcon, fromPlace = non_public, 
                mode = "BICYCLE", cutoffSec = 300)%>% 
  st_transform(crs = MA_state_plane)%>%
  mutate(mode = "bike")
iso_all_modes <- rbind(iso_5min_bike, iso_5min_walk)
```

## Mapping Isochrones

```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(type = "stamenbw", zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), color = "NA", alpha = 0.5) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_manual(values = c("lightgoldenrod1", "orangered3"),
                     name = "Boston Non-Public Schools reachable\nwithin 5 minutes",
                       labels = c("By bike", "By foot")) +
  theme_map() +
  theme(legend.background = element_blank(), legend.position = "left") +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```

## Two different types of visualization
First I will create a graph that uses a cartoon image of a school to signify the points.

```{r}
boston_streets$image <- "school.png"
```

```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(bike))) +
  geom_point() +
    geom_image(aes(image = "school.png"), size = .1, asp = 1.5)+
  scale_x_continuous(name = 
            "Area within a five-minute walking distance of a non-public school\n(square km)",
            breaks = breaks <- seq(10000, 200000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute biking distance of a non-public school\n(square km)",
            breaks = breaks <- seq(0, 1000000, by = 100000),
            labels = breaks / 1000000) +
  theme_pander()
```

Next I'll create a simple violin plot. The data is the same but I think it's easier to accurately analyze the data this way. 

```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(bike))) +
  geom_violin() +  
  geom_point(position = "jitter",
             size = .8,
             alpha = .8) +
  scale_x_continuous(name = 
            "Area within a five-minute walking distance\nof a non-public school\n(square km)",
            breaks = breaks <- seq(10000, 400000, by = 20000),
            labels = breaks / 1000000) +
  scale_y_continuous(name = 
            "Area within a five-minute biking distance\nof a non-public school\n(square km)",
            breaks = breaks <- seq(0, 1000000, by = 100000),
            labels = breaks / 1000000) +
  theme_light()

```

## Acknowledgements

Thank you to Meghan Casey and her code for showing me how to add an image into my ggplot, as well as giving inspiration for pretty isochrone colors!
