---
title: "Vis Assignment 3"
author: "Adriana LH"
date: "9/26/2020"
output: 
 html_document:
    theme: cosmo
    highlight: zenburn
---
```{r load libraries,include=FALSE}
library(sf)
library(tidyverse)
library(ggthemes)
library(ggspatial)
library(units)
library(nngeo)
library(maps)
library(viridis)
```
```{r}
neighborhoods <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/3525b0ee6e6b427f9aab5d0a1d0a1a28_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

precincts <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/2bc185ec9b0b478d8c2f7e720d116863_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

hospitals <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/8f1e5741a4ee4c0f82925d17aab3c002_2.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D", quiet = TRUE)

commcent <- st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/acedd06b39854088b0f2fefccffebf84_0.kml?outSR=%7B%22latestWkid%22%3A3857%2C%22wkid%22%3A102100%7D", quiet = TRUE)
```

## Plotting my data points

I chose to look at Boston neighborhoods and voting precincts (both polygon data) as well as hospitals and community centers in the city (both point data). I was interested in visualizing several types of civic and health resources that are available for Bostonians.

```{r}
ggplot(precincts) +
  geom_sf(color = "yellow", fill = "gray") +
  geom_sf(data = hospitals, size = 1.5) +
  geom_sf(data = commcent, color = "blue", size = 1.5) +
  geom_sf(data = neighborhoods, fill = NA)

```

## Transforming to the Massachusetts State Plane 
```{r}
MA_mainland <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

neighborhoods <- neighborhoods %>% 
  st_transform(MA_mainland)

commcent <- commcent %>% 
  st_transform(MA_mainland)

precincts <- precincts %>%
  st_transform (MA_mainland)

hospitals <- hospitals %>%
  st_transform(MA_mainland)
```

```{r}
ggplot(precincts) +
  geom_sf(color = "yellow", fill = "gray") +
  geom_sf(data = hospitals, size = 1.5) +
  geom_sf(data = commcent, color = "blue", size = 1.5) +
  geom_sf(data = neighborhoods, fill = NA) +
  theme_map() +
  annotation_scale()
```

## Figure 1: Number and proportion of community center points within 1000 meters of a hospital point
```{r}
hospitals_buffer <- st_buffer(hospitals, dist = 1000) %>%
  st_union()

ggplot(hospitals_buffer) + 
  geom_sf() +
  theme_map()
```
```{r}
hospitals_commcent <- commcent[hospitals_buffer,]

ggplot(hospitals_buffer) +
geom_sf() +
  geom_sf(data = hospitals_commcent, color = "blue", size = 2.5) +
  theme_map()
```
```{r}
commcent<- commcent %>%
  st_join(hospitals_commcent) %>%
  mutate(by_hospitals = !is.na(Name.y))

n_hospitals_commcent <- sum(commcent$by_hospitals)

n_hospitals_commcent
```
```{r}
n_commcent <- length(commcent$by_hospitals)
pct_hospitals_commcent <- n_hospitals_commcent/n_commcent

pct_hospitals_commcent
```

There are 14 community centers within 1000 meters of a hospital, which represents 39% of community centers in the city. 

```{r}
left_side  <- st_bbox(hospitals)$xmin
top_side <- st_bbox(hospitals)$ymax

ggplot(neighborhoods) + 
  geom_sf(fill = "gray") +
  geom_sf(data = commcent, size = 2.5, 
          aes(color = by_hospitals)) +
  scale_color_manual(values = c("lightblue", "darkblue"),
name = "Community Centers\nby distance to a hospital", 
labels = c("No community center within 1000 m",
          "Community Center within 1000 m")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_fancy_orienteering()) +
  annotate(geom = "text", x = left_side,
           y = top_side,
           label = paste("Of the ",
                         prettyNum(n_commcent, big.mark = " , "),
                         " community centers in Boston, \n",
                         prettyNum(n_hospitals_commcent, big.mark = ","),
                         " (",
                         prettyNum(100*pct_hospitals_commcent, digits = 0),
                         "%) are within 1000/m of a hospital.",
                         sep = ""),
           hjust = 0, vjust = 0, size =2.3) +
  theme_map() +
  theme(legend.position = "right")
```

## Figure 2: Number and proportion of neighborhoods containing hospitals

```{r}
neighborhoods <- neighborhoods %>%
  mutate(num_hospitals = lengths(st_covers(neighborhoods, hospitals)))

ggplot(neighborhoods) +
  geom_sf(color = "NA", 
          aes(fill = num_hospitals)) +
  scale_fill_viridis_c(name = "Boston Neighborhoods\nby number of hospitals",
                       breaks = breaks <- seq(0, 10, by = 2),
                       labels = paste(prettyNum(breaks, big.mark = "",),
                                      "hospitals")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style =north_arrow_nautical()) +
theme_map() +
theme(legend.position = "right")
```

## Figure 3: Boston neighborhoods by number of community centers

```{r}
neighborhoods <- neighborhoods %>%
  mutate(num_commcent = lengths(st_covers(neighborhoods, commcent)))

ggplot(neighborhoods) +
  geom_sf(color = "black", 
          aes(fill = num_commcent)) +
  scale_fill_viridis_c(name = "Boston Neighborhoods\nby number of Community Centers",
                       breaks = breaks <- seq(0, 5, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = "",),
                                      "Community Centers")) +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style =north_arrow_nautical()) +
theme_map() +
  theme(legend.position = "right")
```

## Figure 4: Overlapping neighborhood and voting precinct polygons

```{r}
neighborhoods <- neighborhoods %>%
  mutate(num_precincts = length(st_overlaps(neighborhoods, precincts))) %>%
  mutate(has_precincts = num_precincts > 0)

num_precincts_neighborhoods <- sum(neighborhoods$has_precincts)
num_precincts_neighborhoods
```

Each neighborhood has a voting precinct. 

```{r}
left_side <- st_bbox(precincts)$xmin
top_side <- st_bbox(precincts)$ymax

ggplot(precincts) +
  geom_sf(fill = "yellow", color = "black") +
  geom_sf(data = neighborhoods, 
          aes(fill = has_precincts)) +
  scale_fill_manual(values = c("yellow", "orange"),
name = "Neighborhoods in Boston\nby presence of voting precincts",
labels = c("Neighborhood without\noverlapping precincts",
           "Neighborhood with\noverlapping precincts")) +
annotation_scale(location = "br") +
  annotate(geom = "text", left_side,
           y= top_side-2000,
           label = paste(num_precincts_neighborhoods, 
                         "of Boston's",
                         length(neighborhoods$Name),
                         "neighborhoods \noverlap with",
                         "precincts."),
                   hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(legend.position = "right")
```

## Figure 5: Average density of hospital points in each neighborhood

```{r}
neighborhoods <- neighborhoods %>%
  mutate(area = set_units(st_area(neighborhoods), km^2)) %>%
  mutate(hospitals_dens = as.numeric(num_hospitals / area))

ggplot(neighborhoods) +
  geom_sf(color = "black",
          aes(fill = hospitals_dens)) +
  scale_fill_viridis_c(name = "Boston Neighborhoods\nby Hospital density",
                       breaks = breaks <- seq(0, 10, by = 1),
                       labels = paste(prettyNum(breaks, big.mark = ","),
                                      "hospitals per square km"),
                       option = "magma") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tl",
                         style = north_arrow_nautical()) +
  theme_map() +
  theme(legend.position = "right")
```

## Figure 6: Mapping the average distance between a community center and a hospital
```{r}
hospitals <- hospitals %>%
  mutate(comm_dist = st_nn(hospitals, commcent, 
                           returnDist = TRUE)$dist) %>%
  mutate(comm_dist = as.numeric(comm_dist))

avg_comm_dist <- mean(hospitals$comm_dist)
avg_comm_dist
```

The average community center is 909 meters away from a hospital. 
```{r}
right_side <- st_bbox(hospitals)$xmax
left_side <- st_bbox(hospitals)$xmin
top_side <- st_bbox (hospitals)$ymax
bottom_side <- st_bbox(hospitals)$ymin

ggplot(precincts) +
  geom_sf(fill= "white", color = "darkgrey") +
  geom_sf(data = hospitals, size = 2,
  aes(color = comm_dist)) +
   coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_color_viridis_c(name = 
                          "Boston hospitals\nby distance to a community center") +
  annotation_scale(location = "br") +
  annotation_north_arrow(location = "tr",
                         style = north_arrow_minimal()) +
  annotate(geom = "text", x = left_side + 850, 
           y = top_side - 1000, 
           label = paste("On average, a hospital\nis ", 
                         prettyNum(avg_comm_dist, digits = 3),
                         " meters from a community center.",
                         sep = ""),
           hjust = 0, vjust = 0, size = 3) +
  theme_map() +
  theme(panel.background = element_rect(fill = "grey"),
        legend.background = element_rect(fill = alpha("white", 0.5), 
                                         color = "gray"))
  
```

